Maximum Entropy Spectral Analysis: a case study

### 2021-05-07 paper/mesa.tex
Edits in paper

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT
\begin{abstract}
The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series. The method relies on a Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR($p$) of order $p$.
A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum}. We characterize our implementation by performing a power spectral density analysis on syntethic data (with known power spectral density) and we compare different criteria for stopping the recursion. Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using both syntethic data and real gravitational wave data from the LIGO-Virgo collaboration.
We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.

@@

The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, astronomy as well as engineering and finance. Of particular interest for several areas of physics is the study of stochastic time series. Among those, a special place is occupied by the so-called \textit{wide-sense} stationary processes. These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation. If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the autocorrelation function 
\begin{equation}
	C(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
\end{equation}
or, equivalently, by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
\begin{equation}\label{eq:psd-autocorrelation}
	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau C(\tau) e^{-i 2 \pi f \tau}\,.
\end{equation}
In some literature, especially in the context of gravitational waves physics, e.g. ~\cite{Finn_1992}, the PSD is introduced as  
\begin{equation}\label{eq:psd-f-definition}
	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
\end{equation}
without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows. The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.

A ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
This approach is equivalent~\cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
\begin{equation}
    \rho_{W} = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\},
\end{equation}
where $\rho$ is the empirical autocorrelation and $M$ is the maximum time lag at which the autocorrelation is computed.
The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.

The choice of a window function is arbitrary and typically is made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of PSD is reached. A high frequency resolution implies high variance and vice-versa.
Besides the window function, Welch's method requires a number of arbitrary choices to be made, such as the number of time slices and the overlap between consecutive slices. All these knobs must be tuned by hand and their choice can dramatically affect the PSD estimation, hence begging the question of what the ``real" PSD is.

@@

An appealing alternative approach, based on the Maximum Entropy principle \cite{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by Burg \cite{burg1975maximum}. Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series. Furthermore, it embeds the PSD estimation problem into an elegant theoretical framework and makes minimal assumptions on the nature of the data.
Lastly and most importantly, it provides a robust link between spectral density estimation and the field of autoregressive processes. This provide a natural and simple machinery to forecast a time series, thus predicting future observations based on previous ones.

In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg's algorithm and the link between Burg's algorithm and autoregressive process.
Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problems posed by the Welch's algorithm (or other similar methods).
To facilitate this goal, we present and describe a new implementation, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
Available at link: \url{https://pypi.org/project/memspectrum/}.
}.
We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests on simulated and real data.
We also compare our results with those of spectral analysis carried out with the standard Welch's method.
In order to apply our model on a realistic setting, we analyse some time series of broad interest in the scientific community.

Our paper is organized as follows: we being by briefly reviewing the theoretical foundations of the maximum entropy principle in Sec.~\ref{sec:foundations}. Sec.~\ref{sec:validation} presents the validation of Burg's method as well as of our implementation on simulated data. In Sec.~\ref{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.

@@

Before introducing MAXENT principle, we will define through some simple examples the two core concepts of the problem and the role they play: the 'evidence' and the 'information'.
Let us start with the 'information' (or entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study. For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information content. Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$), as in this case, a communication is not even needed.  
Shannon \cite{Shannon} proposed the quantity
\begin{equation}\label{eq:information}
    I = \log_2 \frac{1}{p(x)}
\end{equation}
to measure the quantity of information brought by an outcome $x$ with probability $p(x)$. It is additive quantity as well as monotonically decreasing as a function of $p \in [0, 1]$: the more uncertain the outcome, the higher the information it brings.

We can generalize the definition of information in the case where two different outcomes $E_1, E_2$, with given probabilities $P_1$ and $P_2$, are possible.
To gain some intuition on the problem, we ask ourselves which are the probability assignments that makes the outcome more uncertain (i.e. maximize the information).
If $P_1$ and $P_2$ are largely different, for instance $P_1 = 0.999$ and $P_2 = 0.001$, we are allowed to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome. The information content will be very low.
On the other hand, most unpredictable experiment happens when 
\begin{equation}\nonumber
    P_1 = P_2 = \frac{1}{2}:
\end{equation}
this describes a situation of `maximum ignorance': the information content of this system must be high.
Any generalization of eq.~\eqref{eq:information}, must then have its maximum when $ P_1 = P_2$.
For $N$ events, system with the highest possible information content is when:
\begin{equation}\nonumber
    P_1 = \hdots = P_N = \frac{1}{N}:
\end{equation}

@@

Of particular importance is the case in which, in addition to the mean, also the variance is known. MAXENT leads to the Gaussian distribution. 
We retain this particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
Indeed, the Gaussian distribution is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy. For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process. In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy.

@@

In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
would reduce to a single Fourier transform.
However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
is impossible.
Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
}, so that only a few values for the autocorrelation function can actually be computed reliably.
This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function? MAXENT can guide us in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method. The latter assumes that the unknown values of the autocorrelation are $0$. Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.

As in the previous examples, one needs to set up a variational problem where the entropy eq.~\eqref{eq:entropy} is maximized 
subject to some problem-specific constraints. 
In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).

Before doing so, there is a technical problem to solve: the definition of entropy depends on a probability distribution, not on the PSD.
It can be shown, e.g.~\cite{AblesMESA, Bartlett}, that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
$$T(f) = \frac{\tilde{y}(f)}{\tilde{x}(f)}$$
where $\tilde{x}(f)$ denotes the Fourier transform of $x_t$ (and similarly for $y_t$)
}.
The difference in entropy between the input and the output time series (i.e. the entropy gain) obtained by such filter applied on white noise is:
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
\end{equation}
where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximising Eq.~\eqref{eq:entropy_continuos}.

@@

Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
This approach on PSD computation provide a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.
Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
The expression was first found by Burg~\cite{burg1975maximum}:
\begin{equation}\label{eq:MESApsd}
    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
\end{equation}
where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1.
The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
The coefficients $P_N$ and $a_s (s > 0)$ are to be determined by an iterative process (called Burg's algorithm). The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:MESA_solution}.

@@

The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
of autoregressive processes (AR)~\cite{doi:10.1029/RG013i001p00183}.
An autoregressive stationary process of order $p$ (AR($p$)) is a time series whose values satisfy the following expression: 
\begin{equation} \label{eq:AR_p}
    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_p x_{t - p} = \nu_t
\end{equation}
where $b_1, \ldots, b_p$ are real coefficients and $\nu_t$ is white noise with a given variance $\sigma^2$.
Thus, an AR($p$) process models the dependence of the value of the process at time $t$ on every past $p$ observations, 
thus being potentially able to model complex autocorrelation structures within observations.

@@

Eqs.~\eqref{eq:ARspectrum} and ~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD also computes the coefficients of the AR 
process that models the observed process.

\subsection{Forecasting} \label{sec:forecasting}
The link between MESA and AR processes is of particular interest. Given the solution to Burg's recursion to determine the $a_k$, see Appendix~\ref{sec:appendix}, we automatically obtain the coefficients of the equivalent AR process, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
Indeed, for an AR($p$) process the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation has the form:
\begin{align}\label{eq:p_forecast}
	p&(x_t|x_{t-1}, \ldots , x_{t-p}) \nonumber\\
	&= \frac{1}{\sigma\sqrt{2\pi}} \exp\left[-\frac{1}{2} \left(\frac{x_t - \sum_{i = 1}^p b_i x_{t-i}}{\sigma}\right)^2\right]\,.
\end{align}
The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computations.

\section{Validation of the model}\label{sec:validation}
It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}). The number $M$ of such coefficients is the same as the number of time steps considered in the computation of the autocorrelation $\bar{r}_m$. In an ideal scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered). However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $m$, as the estimation for  $\bar{r}_m$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{M - m}\right)^{-1}$.

@@

The reminder of this section is devoted to an extensive study on how to make such choice.
In Section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the 
algorithm is able to reproduce a known PSD.
The basic idea is to validate, as the autoregressive order considered increase, the performance of the algorithm results 
by measuring the loss function and pick, among the orders the one that yields better results.
Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with given analytical PSD, Sec.~\ref{sec:arp_validation}.
In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on real data from the LIGO-Virgo gravitational waves observatories.

@@

\begin{itemize}
\item \textbf{Final prediction Error} 
The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. He proposed that $m$ should be chosen as the 
length that minimises the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
\begin{equation}
    FPE(m) = \mathbb{E}\left[ \left((x_t - \hat x_t) ^ 2\right) \right]
\end{equation}
with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
Minimizing FPE is equivalent to minimizing the quantity: 
\begin{equation}
    \mathcal{L}_{\rm FPE}(m) = P_{m} \frac{N + m + 1}{N - m - 1}
\end{equation}
with $P_m$ being the estimated noise variance at order $m$, see Eq.~\eqref{eq:errorFilter1}. In the $N \to \infty$ limit, 
remembering $m_{max} \sim 2N / \log(2N)$, Akaike's loss function is equivalent to the minimization of the variance $P_m$ of the white noise of the underlying $AR(p)$ model.

@@

\item \textbf{Optimum Bayes Decision rule} 
The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
time series for the process that has to be described as an autoregressive process of order $m$ ($AR(m)$), with $m$ to be determined. The OBD is obtained choosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$. Each
hypothesis is uniquely determined from both the length and the values of the filter under study. Choosing a Likelihood of the form \eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e. maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:

>>>>

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT
\begin{abstract}
The Maximum Entropy Spectral Analysis (MESA) method, developed by Burg, provides a powerful tool to perform spectral estimation of a time-series. The method relies on a Jaynes' maximum entropy principle and provides the means of inferring the spectrum of a stochastic process in terms of the coefficients of some autoregressive process AR($p$) of order $p$.
A closed form recursive solution provides an estimate of the autoregressive coefficients as well as of the order $p$ of the process.
We provide a ready-to-use implementation of the algorithm in the form of a python package \texttt{memspectrum}. We characterize our implementation by performing a power spectral density analysis on syntethic data (with known power spectral density) and we compare different criteria for stopping the recursion. Furthermore, we compare the performance of our code with the ubiquitous Welch algorithm, using syntethic data generated from the released spectrum by the LIGO-Virgo collaboration.
We find that, when compared to Welch's method, Burg's method provides a power spectral density (PSD) estimation with a 
systematically lower variance and bias. This is particularly manifest in the case of a little number of data points, making Burg's method most suitable to work in this regime.

@@

The study of the properties of stochastic processes is a crucial task in many fields of physics, astronomy, quantitative biology, as well as engineering and finance. Among those, a special place is occupied by the so-called \textit{wide-sense} stationary processes. These are stochastic processes that display an invariance of their statistical properties, such as their two-point autocovariance function, with respect to time translation. If $x(t)$ is a wide-sense stationary process, it is completely determined by the knowledge of the autocorrelation function 
\begin{equation}
	C(\tau) = \mathbf{E}[x_t \cdot x_{t+\tau}]
\end{equation}
or, equivalently, by the knowledge of their \emph{power spectral density} (PSD) $S(f)$. Thanks to the Wiener-Khinchin theorem the two are related by a Fourier transform: 
\begin{equation}\label{eq:psd-autocorrelation}
	S(f) = \int_{-\infty}^{\infty} \textrm{d}\tau C(\tau) e^{-i 2 \pi f \tau}\,.
\end{equation}
In some literature, especially in the context of gravitational waves physics, e.g. ~\cite{Finn_1992}, the PSD is introduced as  
\begin{equation}\label{eq:psd-f-definition}
	\mathbf{E}[\tilde{x}(f) \cdot \tilde{x}(f^\prime)] = S(f) \delta(f-f^\prime)
\end{equation}
without highlighting its connection with the time structure of the process itself, thus masking some important properties that will explore further in what follows. The latter definition in~\eqref{eq:psd-f-definition} gives, however, i) a straightforward interpretation of the PSD: it measures how much signal ``power" is located in each frequency; ii) an operative way of estimating it for an unknown process.

An ubiquitous method for such computation is due to Welch \cite{Welch1967} and it based on eqs.(\ref{eq:psd-autocorrelation}-\ref{eq:psd-f-definition}).
The PSD is obtained by slicing the observed realisation $x(t_1),\ldots,x(t_n)$ of the process $x(t)$ into many window-corrected batches and averaging the squared moduli of their Fourier transforms.
This approach is equivalent~\cite{Lomb,Scargle} to taking the Fourier Transform of the windowed sample autocorrelation $\rho_W$, written as
\begin{equation}
    \rho_{W} = \left\{W_0\rho_0,W_{\pm 1}\rho_{\pm 1}, \dots, W_{\pm M}\rho_{\pm M}, 0, 0, \dots \right\},
\end{equation}
where $\rho$ is the empirical autocorrelation and $M$ is the maximum time lag at which the autocorrelation is computed.
The sequence $W$ is a window function that can be chosen in several different way, each choice presenting advantages and disadvantages for the final estimate of the PSD.

The choice of a window function is arbitrary and typically is made by trial and error, until a satisfactory compromise between variance and resolution of the estimate of PSD is reached. A high frequency resolution implies high variance and vice-versa.
Besides the window function, Welch's method requires a number of arbitrary choices to be made, such as the number of time slices and the overlap between consecutive slices. All these knobs must be tuned by hand and their choice can dramatically affect the PSD estimation, hence begging the question of what the ``best" PSD estimate is.

@@

An appealing alternative approach, based on the Maximum Entropy principle \cite{JaynesArticle,jaynes2003ptl, Jaynes_MAXENT}, has been put forward by Burg \cite{burg1975maximum}. Being rooted on solid theoretical foundations, we will see that Burg's method, unlike Welch's, does not require any preprocessing of the data and requires very little tuning of the algorithm parameters, since it provides an iterative closed form expression for the spectrum of a stochastic stationary time series. Furthermore, it embeds the PSD estimation problem into an elegant theoretical framework and makes minimal assumptions on the nature of the data.
Lastly and most importantly, it provides a robust link between spectral density estimation and the field of autoregressive processes. This provides a natural and simple machinery to forecast a time series, thus predicting future observations based on previous ones.

In this paper, we discuss the details of the Maximum entropy principle, its application to the problem of PSD estimation with Burg's algorithm and the link between Burg's algorithm and autoregressive process.
Our goal is to bring (again) to public attention Maximum Entropy Spectral analysis, in the hope that it will be widely employed as a way out to the many problems posed by the Welch's algorithm (or other similar methods).
To facilitate this goal, we present and describe a new code, \texttt{memspectrum}, that provides a robust and easy-to-use python implementation of the algorithm\footnote{
It is available at link: \url{https://pypi.org/project/memspectrum/}.
}.
We provide a thorough assessment of the performance of our code and we validate our results performing a number of tests on simulated and real data.
We also compare our results with those of spectral analysis carried out with the standard Welch's method.
In order to apply our model on a realistic setting, we analyse some time series of broad interest in the scientific community.

Our paper is organized as follows: we being by briefly reviewing the theoretical foundations of the maximum entropy principle in Sec.~\ref{sec:foundations}. Sec.~\ref{sec:validation} presents the validation of Burg's method as well as of our implementation on simulated data. In Sec.~\ref{sec:Welch_comparison} we compare the results from \texttt{memspectrum} with the Welch method; Sec.~\ref{sec:applications} presents a few applications to real time series and, finally, we conclude with a discussion in Sec.~\ref{sec:conclusion}.

@@

Before introducing MAXENT principle, we will define through some simple examples the two core concepts of the problem and the role they play: the `evidence' and the `information'.
Let us start with the `information' (or entropy): it is a measure of the degree of uncertainty on the outcomes of some experiment and specifies the length of the message necessary to provide a full description of the system under study. For instance, consider a perfectly sinusoidal signal: knowledge of amplitude, frequency and phase are sufficient to fully reproduce it: it has a low information content. Even less information is required if we are studying a system whose outcome is certain (has probability $p = 1$), as in this case, a communication is not even needed.  
Shannon \cite{Shannon} proposed the quantity
\begin{equation}\label{eq:information}
    I = \log_2 \frac{1}{p(x)}
\end{equation}
to measure the quantity of information brought by an outcome $x$ with probability $p(x)$. It is additive quantity as well as monotonically decreasing as a function of $p \in [0, 1]$: the more uncertain the outcome, the higher the information it brings.

We can generalize the definition of information in the case where two different outcomes $E_1, E_2$, with given probabilities $P_1$ and $P_2$, are possible.
To gain some intuition on the problem, we ask ourselves which are the probability assignments that makes the outcome more uncertain (i.e. maximize the information).
If $P_1$ and $P_2$ are largely different, for instance $P_1 = 0.999$ and $P_2 = 0.001$, we are allowed to believe that event $E_1$ will occur almost certainly, considering $E_2$ to be a very implausible outcome. The information content will be very low.
On the other hand, most unpredictable experiment happens when 
\begin{equation}\nonumber
    P_1 = P_2 = \frac{1}{2}:
\end{equation}
this describes a situation of `maximum ignorance' and the information content of such system must be high.
Any generalization of eq.~\eqref{eq:information}, must then have its maximum when $ P_1 = P_2$.
For $N$ events, system with the highest possible information content is when:
\begin{equation}\nonumber
    P_1 = \hdots = P_N = \frac{1}{N}:
\end{equation}

@@

Of particular importance is the case in which, in addition to the mean, also the variance is known: MAXENT leads to the Gaussian distribution. 
We retain this particularly interesting from the foundational point of view, since it provides a deeper insight into the ubiquitous Gaussian distribution.
Indeed, it is not only the limit distribution provided by the central limit theorem for finite variance processes but it is also the distribution that maximizes the entropy. For this reason, appealing to MAXENT principle, it is the correct assignment if mean and covariance are the only quantities that fully define our process. In some sense, we can interpret the central limit theorem as the natural `statistical' evolution toward a configuration that maximizes entropy.

@@

In principle, if the autocorrelation was known exactly (i.e. at every time $\tau \in (-\infty,+\infty)$), the computation of the PSD 
would reduce to a single Fourier transform.
However, in any realistic setting, we are dealing with a finite number of samples $N$ from the process, hence such computation
is impossible.
Moreover, the error $\sigma_k$ in the estimate of the autocorrelation after $k$ steps increases as $\sigma \sim 1/\sqrt{N - k}$\footnote{
This is clearly understood: when computing the autocorrelation at order $k$, only $N-k$ examples of the product $x_t x_{t+k}$ are available and the variance of the average value goes as the inverse of the square root of the points considered.
}, so that only few values for the autocorrelation function can actually be computed reliably.
This bring us the core of the problem: how to give an estimate from partial (and noisy) knowledge of the autocorrelation function? MAXENT can guide us in this task without any a priori assumptions on the unavailable data\footnote{Indeed this is the largest difference with the most common Welch method. The latter assumes that the unknown values of the autocorrelation are $0$. Clearly, this assumption is unjustified and MAXENT is a good way to drop this assumption.}.

As in the previous examples, one needs to set up a variational problem where the entropy eq.~\eqref{eq:entropy_continuos} is maximized 
subject to some problem-specific constraints. 
In our case, they are i) the PSD estimate has to be non-negative; ii) its Fourier transform has to match the sample autocorrelation (wherever an estimate of this is available).

Before doing so, there is a technical problem to solve: the definition of entropy depends on a probability distribution, not on the PSD.
It can be shown, e.g.~\cite{AblesMESA, Bartlett}, that the variational problem can be formulated in terms of the power spectral density $S(f)$ alone by considering our signal as the result of the filtering a white noise process using a filter with transfer function $T(f)$ equal to $S(f)$\footnote{
A filter with transfer function $T(f)$ takes in input a time series $x_t$ and outputs a times series $y_t$ such that:
$$T(f) = \frac{\tilde{y}(f)}{\tilde{x}(f)}$$
where $\tilde{x}(f)$ denotes the Fourier transform of $x_t$ (and similarly for $y_t$)
}.
The difference in entropy between the input and the output time series (i.e. the entropy gain) obtained by such filter applied on white noise is:
\begin{equation}\label{eq:EntropyGain}
    \Delta H = \int_{- Ny}^{Ny}\log S(f) df\,.
\end{equation}
where $\Delta t$ is sampling rate and $Ny \equiv \frac{1}{2 \Delta t}$  is the Nyquist frequency.
Thus maximising Eq.~\eqref{eq:EntropyGain} is equivalent to maximise Eq.~\eqref{eq:entropy_continuos}.

@@

Thus, by maximizing Eq.~\eqref{eq:EntropyGain} with constraints in eq.~\eqref{eq:MaxConstraint}, we can give an estimate of the spectrum given an empirical time series.
This approach on PSD computation provides a result consistent with the epirical autocorrelation function whenever this is available and, at the same time, it does not make any assumption for the unavailable estimates for the autocorelation at large time lags.

Most importantly, the variational problem admits a closed-form analytical expression for $S(f)$.
The expression was first found by Burg~\cite{burg1975maximum}:
\begin{equation}\label{eq:MESApsd}
    S(f) = \frac{P_N \Delta t}{\left(\sum_{s=0}^N a_s z^s\right)\left(\sum_{s = 0}^N a^*_s z^{-s}\right)}, 
\end{equation}
where $\Delta t$ is the sampling interval of the time series, $z=\exp{(2\pi i f\Delta t)}$, $a_0$ = 1.
The vector obtained as $(1, a_1, \dots, a_N)$ is also known as the \textit{prediction error filter}.
The coefficients $P_N$ and $a_s (s > 0)$ are to be determined by an iterative process (called Burg's algorithm). The details of the derivation and the actual form for the coefficients $a_s$ can be found in appendix \ref{sec:appendix}.

@@

The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
of autoregressive processes (AR)~\cite{doi:10.1029/RG013i001p00183}.
An autoregressive stationary process of order $p$, AR($p$), is a time series whose values satisfy the following expression: 
\begin{equation} \label{eq:AR_p}
    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_p x_{t - p} = \nu_t
\end{equation}
where $b_1, \ldots, b_p$ are real coefficients and $\nu_t$ is white noise with a given variance $\sigma^2$.
Thus, an AR($p$) process models the dependence of the value of the process at time $t$ on every past $p$ observations, 
thus being potentially able to model complex autocorrelation structures within observations.

@@

Eqs.~\eqref{eq:ARspectrum} and ~\eqref{eq:MESApsd} are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD models the observed times series as an AR process and provides a {\it fit} for the autoregressive coefficients.
Furthermore, as a consequence of Wold's theorem, there is the theoretical guarantee that every stationary time series can be modelled faithfully by the MAXENT.

\subsection{Forecasting} \label{sec:forecasting}
The link between MESA and AR processes is of particular interest. Given the solution to Burg's recursion to determine the $a_k$, we automatically obtain the coefficients of the equivalent AR process, hence we are able to exploit Eq.~\ref{eq:AR_p} to perform \emph{forecasting}, thus providing plausible future observations, conditioned on the observed data.
Indeed, for an AR($p$) process the conditional probability $p(x_t|x_{t-1}, \ldots , x_{t-p})$ of the observation at time $t$ with respect to the past $p$ observation has the form:
\begin{align}\label{eq:p_forecast}
	p&(x_t|x_{t-1}, \ldots , x_{t-p}) \nonumber\\
	&= \frac{1}{\sigma\sqrt{2\pi}} \exp\left[-\frac{1}{2} \left(\frac{x_t - \sum_{i = 1}^p b_i x_{t-i}}{\sigma}\right)^2\right]\,.
\end{align}
The interpretation of Eq.~(\ref{eq:p_forecast}) is straightforward: $x_t$ follows a Gaussian distribution with a fixed variance an a mean value $m_t = \sum_{i = 1}^p b_i x_{t-i}$ computed from past observations.
Eq.~(\ref{eq:p_forecast}) provides then a well defined probability framework for predicting future observations: this is a very useful feature of MESA, that does not have an equivalent in any other spectral analysis computation methods.

\section{Validation of the model}\label{sec:validation}
It is now clear that MESA provides a recursive formula for computing the coefficients $a_k$ in Eq.~(\ref{eq:MESApsd}). The number $M$ of such coefficients is
equivalent to the maximum order of the autocorrelation $\bar{r}_m$ considered. In an ideal scenario, this would be equal to the number of points the autocorrelation is computed at (equivalent to the length of data considered). However, the computation of high order coefficients of the autocorrelation is unstable and for high enough $m$, as the estimation for  $\bar{r}_m$ shows a very high variance, broadly scaling as $\sim \left(\sqrt{M - m}\right)^{-1}$.

@@

The reminder of this section is devoted to an extensive study on how to make such choice.
In Section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the 
algorithm is able to reproduce a known PSD.
The basic idea is to validate, as the autoregressive order considered increase, the performance of the algorithm results 
by measuring the loss function and pick, among the orders the one that yields better results.
Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with an analytical gaussian PSD, sec.~\ref{sec:arp_validation}.
In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on syntetic data generated from the analytical spectrum released by the LIGO-Virgo collaboration.

@@

\begin{itemize}
\item \textbf{Final prediction Error} 
The first criterion is due to Akaike~\cite{Akaike1970StatisticalPI}. It was proposed that $m$ should be chosen as the 
length that minimises the error when the filter is used as a predictor, the \emph{final prediction} error (FPE): 
\begin{equation}
    FPE(m) = \mathbb{E}\left[ \left((x_t - \hat x_t) ^ 2\right) \right]
\end{equation}
with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
Minimizing FPE is equivalent to minimizing the quantity: 
\begin{equation}
    \mathcal{L}_{\rm FPE}(m) = P_{m} \frac{N + m + 1}{N - m - 1}
\end{equation}
with $P_m$ being the estimated noise variance at order $m$, see Eq.~\eqref{eq:errorFilter1}. In the $N \to \infty$ limit, 
remembering $m_{max} \sim 2N / \log(2N)$, Akaike's loss function is equivalent to the minimization of the variance $P_m$ of the white noise of the underlying $AR(p)$ model.

@@

\item \textbf{Optimum Bayes Decision rule} 
The last criterion we will consider is the Optimum Bayes Decision rule (OBD)\cite{doi:10.1029/WR018i004p01097}. Let $x(t)$ be the observed 
time series for the process that has to be described as an $AR(m)$ process, with $m$ to be determined. The OBD is obtained choosing between M different hypothesis $H_i, i = 1, \dots, M$ maximizing their posterior distribution $P(H_i\vert x(t))$. Each
hypothesis is uniquely determined from both the length and the values of the filter under study. Choosing a Likelihood of the form eq.~\eqref{eq:p_forecast} and uniform priors for both the coefficients and the hypotheses, Rao has shown that choosing the minimum for $-log(P(H_i \vert x(t))$, i.e. maximizing the posterior for $H_i$, is asymptotically equivalent to the minimization of:
